---
title: "Boring Report"
author: "Ryan Johnson"
---

```{r}
library(tidyverse)
library(pins)
```

## **Introduction**

Posit Package Manager (PPM) is a tool for organizing, centralizing, and distributing R and Python packages for your team. It can also be used to host internally developed R/Python packages. PPM comes with an experimental API which can be used to gather metrics regarding package downloads. In this report, we'll query the API within [Posit's Public Package Manager](https://packagemanager.posit.co/client/#/) to gather download statistics for three common R packages:

-   [`ggplot2`](https://ggplot2.tidyverse.org/index.html) - system for declaratively creating graphics, based on [The Grammar of Graphics](https://www.amazon.com/Grammar-Graphics-Statistics-Computing/dp/0387245448/ref=as_li_ss_tl).

-   [`dplyr`](https://dplyr.tidyverse.org/) - grammar of data manipulation, providing a consistent set of verbs that help you solve the most common data manipulation challenges.

-   [`shiny`](https://shiny.posit.co/) - build interactive web applications (apps) straight from R.

## Package Download Metrics

```{r}
# Connect to Posit Connect 
board <- board_connect(auth = "envvar")

# Read pin (total downloads)
package_counts_30 <- pin_read(board, "ryan/package_counts_30")

# Read pin (download per day)
package_daily_counts_30 <- pin_read(board, "ryan/package_daily_counts_30")
```

```{r}
# Create gt table
package_counts_30 
```

```{r}
# Create plot
ggplot(package_daily_counts_30, aes(x = as_date(date), y = downloads, color = package)) + 
  geom_smooth(linewidth = 1.3)
```
